Qualifying Report

### 2019-01-30 Source/QualifyingReport/QualifyingReport.tex
stats and QR update

I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H. W. Bush (GHWB) and 25 press conferences conducted by Donald J. Trump (DJT).  I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above). For this collection of documents, I generated a number of features which looked at a number of different aspects of each document. These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings. These findings were. number of unique words, non-specific nouns and fillers and low imageability (LI) verbs. Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image. I compared the trends described in the transcripts of RR and GHWB, but also included DT. Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days). However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days). It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age). DJT as with GHWB has no known diagnosis of AD. I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project. The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.

@@

\section{Results}
	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
		& RR & GHWB & DJT \\ \hline
		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline

	\end{tabular}
	\caption{\label{tab:table-name}NLTK tasks and functionality}
	\end{center} 
\end{table}
\section{Discussion}
\section{Conclusion}

>>>>

I took 46 transcripts of Ronald Reagan’s (RR) press conferences from 1981 to 1988 and compared them with 134 press conferences by George H. W. Bush (GHWB) and 29 press conferences conducted by Donald J. Trump (DJT).  I analyzed transcripts for lexical features shown to change longitudinally with dementia  (for a comprehensive review of these, see the literature review above). For this collection of documents, I generated a number of features which looked at a number of different aspects of each document. These features encompassed, word level, sentence level and document level features and included a number of features contained in the study by Berisha and Liss with the aim of replicating and extending on their findings. These findings were. number of unique words, non-specific nouns and fillers and low imageability (LI) verbs. Imageability is characterized, according to Berisha and Liss, as the ease with which a term gives rise to a sensory .mental image. I compared the trends described in the transcripts of RR and GHWB, but also included DJT. Berisha and Liss originally made the comparison as it GHWB (GHWB - age at the start of presidency - 64 years, 222 days) was the closest match to RR in terms of age (RR  - age at the start of presidency, 69 yeaars and 349 days). However, with the inauguration of Trump, he now is the closest comparible president in terms of age (DJT - age at the start of presidency - 70 years, 220 days). It would be interesting to look at a comparison of RR and DJT to see whether the comparisons made by Berisha and Liss hold true with this more appropriate match (in terms of age). DJT as with GHWB has no known diagnosis of AD. I used the press conference transcripts in the American Presidency Project (APP) archive as a data source for this project. The APP is a comprehensive and organized searchable database of presidential documents, including transcripts of speeches, transcripts of news conferences, and other public documents.

@@

We calculating the following features for each transcript in turn using the NLTK (see section 2 for a description) \cite{Bird2009} and Python. 
\subsubsection{Measures of lexical variation}
% CITE Richards 1986 Type/Token Ratios: What do they really tell us?
We constructed two features of lexical variation. Firstly we looked at the number of unique words. To do this we were able to split each transcript into individual words and changed them to lowercase using NLTK and were then count the number of unique words that appeared in each transcript. We also used the TTR formula (see section 2 for a description) for a feature that measures lexical diversity independent of sample size \cite{}. 
\subsubsection{Fillers, Non-Specific Nouns and LI Verbs}
For these features, we counted the number of occurrences for different categories (see table for list of categories tracked and the words counted.

\begin{table}
	\begin{center}
	\begin{tabular}{ | p{3cm} | p{6cm} |}
		\hline
		Category & Words \\ \hline
		Fillers & \textit{"well", "so", "basically", "actually", "literally", "um", "ah"} \\ \hline
		Non Specific Nouns & \textit{"something", "anything", "thing", "everything"} \\ \hline
		LI Verbs & \textit{"be", "come", "do", "get", "give", go", "know", "look", "make", "see", "tell", "think", "want"} \\ \hline

	\end{tabular}
	\caption{\label{tab:table-name}Categories and Words Counted}
	\end{center} 
\end{table}

\subsubsection{Usage of parts of speech}
This section involves using a Part of Speech tagger (PoS) which analyses a sentence and assigns a 'tag' to each word based on the function the word has in a sentence. At a basic level this can be divided into the eight defined parts of speech: 'nouns', 'pronouns', verbs', 'adjectives', 'adverbs', 'conjunctions', 'prepositions' and 'interjections' but can be further subcategorised. We used the PoS tagger built into NLTK to tag each transcript in turn and used these the counts from each of these eight categories in our analysis. In addition to frequency counts we also normalised these features by dividing the frequency count by the number of words in the document to take into account transcript length.

\section{Results}
\begin{table}
	\begin{center}
	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
		\hline
		& RR & GHWB & DJT \\ \hline
		Total Words & 3423.91 (416.42) & 2607.72 (1210.38) & 1848.65 (1549.38) \\ \hline
		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
	\end{tabular}
	\caption{\label{tab:table-name}Means and Standard Deviations of important features}
	\end{center} 
\end{table}

\begin{table}
	\begin{center}
	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
		\hline
		& RR & GHWB & DJT \\ \hline
		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline

	\end{tabular}
	\caption{\label{tab:table-name}RR T-tests vs GWB and DJT}
	\end{center} 
\end{table}

\begin{table}
	\begin{center}
	\begin{tabular}{ | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
		\hline
		& RR & GHWB & DJT \\ \hline
		Unique Words & 894.13 (85.15) & 667.76 (218.67) & 481.82 (221.29) \\ \hline
		Non Specific Nouns & 12.72 (4.63) & 6.78 (4.32) & 7.41 (8.75) \\ \hline
		LI Verbs & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Nouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Verbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Adjectives Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Adverbs Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Pronouns Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Conjunctions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Prepositions Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline
		Interjections Normalised & 124.22 (17.89 )& 103.45 (51.87) & 84.48 \\ \hline

	\end{tabular}
	\caption{\label{tab:table-name}Pearson Correlations for Features}
	\end{center} 
\end{table}

\section{Discussion}
% Cite Garrard 2005 -The effects of very early Alzheimer's disease on the characteristics of writing by a renowned author.

Our research supports the findings of Berisha and Liss with regards to unique words, non-specific nouns and low-imageability words\cite{Berisha2015}.
\par 
A criticism of Berisha and Liss's work is the problems they had with normalising the transcripts in terms of length. This was also a problem in the work of Garrard et al. Whilst it is important to control for outliers, there are other ways in which we can control for length of sample. In this study, we explored features such as TTR, and also normalised versions of the frequency statistics we used, and also dividing the frequency counts by 100 which and these showed similar significant results.

 In addition, when adding DJT as an additional control to compare to and despite a comparatively smaller sample per press conference held, their findings also hold true. 
\par 
It's worth noting that DJT's conferences are, with the exception of one press conference, held jointly with other people and therefore the amount of speech produced per conference is a lot less. This finding is, in itself notable as it illustrates the problems with amounts of speech. This problem could potentially be solved with 
\par 
There are limitations of this research. Whilst in terms of age, DJT is certainly more suitable as a control to match with RR, in some ways they held very different styles of press conferences in that RR preferred to do solo press conferences and DJT has shown a preference for doing joint press conferences which have an impact on the amount of language produced. Another limitation is the way in which I normalised the features. 
\par 
With further work, we will not have the luxury of having multiple samples and so it would be worth exploring how the quality of these predictions might lessen when faced with considerably fewer samples.
\section{Conclusion}
The results of this work show that we can track a person's use of language through time in a number of ways, and that
